Flesh Simulation in the Field of Animation

### 2020-04-04 LaTeX-Master-Vorlage/chapter/Background.tex
working on paper

\chapter{Background} \label{c:Background}
As in the paper \textit{\acrshort{snh}}, my goal is to animate human-like characters. More precisely, the focus is on the behaviour of the character's flesh. This chapter serves as an introduction into the mathematical as well as physical background, needed to understand the upcoming calculations and conclusions. At the beginning of this chapter I will define the notation and convention used throughout this whole thesis. Next, I will give insights in the mathematics used and present some of the concepts used in continuum mechanics. However, I will not include each proof explicitly, as there are already many good resources for an interested reader.

@@

The \acrlong{svd} (\acrshort{svd}) will play an important role in the formulation of the deformation gradient. It represents the best possible approximation of a given matrix by a matrix of low rank. This approximation can be looked at as a compression of the data given (\cite{LiesenMehrmann2015}, S. 295).

@@

This definition and theorem were taken from \cite{ford2014numerical} (p.113, p.300).

\subsection{Polar Decomposition}

@@

This theorem was taken from \cite{Spencer1980} (p.12).

@@

We can imagine, that we map each particle of a chosen object from its rest state to a deformed one. Each particle $X$ of a body can be characterized by a vector $\mathbf{x}$ containing its positional coordinates at time $t_0=0$. This is the reference configuration. If the particles are displaced after a time $t$, we can describe their new coordinate vector with 
\[
	\check{\mathbf{x}} = \phi (\mathbf{x}, t).
\]
If we consider, for example a \textit{Translation}, every particle is displaced by the same distance and direction. Hence, we can calculate $\check{\mathbf{x}}$ with
\[
	\check{\mathbf{x}} = \mathbf{x} + \mathbf{c}(t)
\]
where $\mathbf{c}$ is a vector that only depends on $t$ (\cite{Spencer1980}, p.63).

@@

\subsection{Deformation Gradient}
The next quantity, I will use in the upcoming chapters, is the deformation gradient \textbf{F}. With its help, we can calculate the change in volume and length of an object during a deformation. We can obtain \textbf{F} by taking the gradient of the function $\phi$ discussed in the previous section.

\begin{wrapfigure}[11]{r}{5.5cm}
\centering
	\includegraphics[width=5.5cm]{resources/stretch_plot.png}
	\caption[Stretching of a rectangle]{Stretching of a rectangle}
	\label{stretch:1}
\end{wrapfigure}
As an example, I am taking a simple deformation in 2D. Consider the coloured area in Fig. \ref{stretch:1}. We can stretch this area in the following way:
\begin{align*}
	\check{x} &= 1.5x + 0y		\\
	\check{y} &= 0x + 2y	
\end{align*}

The resulting deformation gradient can then be obtained by
\[
	\mathbf{F} = \begin{bmatrix} \frac{\partial [1.5x + 0y]}{\partial x} & \frac{\partial [1.5x + 0y]}{\partial y} \\ \frac{\partial [0x + 2y]}{\partial x} & \frac{\partial [0x + 2y]}{\partial y} \end{bmatrix} = \begin{bmatrix} 1.5 & 0.0 \\ 0.0 & 2.0 \end{bmatrix}.
\]
If \textbf{F} is equal to the identity matrix \textbf{I}, there is no deformation present. This would be the case for rigid body displacements. Since we will be handling deformation in the 3D space, the deformation gradient will have the form of a $3 \times 3$ matrix: 
\begin{equation}\label{eq:deformation_gradient}
\textbf{F} = \left[ \,f_0\, \bigg| \,f_1\, \bigg| \,f_2\, \right] = \begin{bmatrix} f_0 & f_3 & f_6 \\ f_1 & f_4 & f_7 \\ f_2 & f_5 & f_8 \end{bmatrix}.
\end{equation}
\textbf{F} can be factorized and used to calculate other quantities. The next few sections show these factorizations and calculations.

@@

\subsubsection{Cauchy-Green}
The right Cauchy-Green tensor \textbf{C} can be calculated by 
\begin{equation}\label{CG_DG}
	\mathbf{C} = \mathbf{F^\intercal F}
\end{equation}
and is a $3 \times 3$ matrix for deformations in the 3D space. Using \textbf{C}, we can calculate the first right Cauchy-Green invariant $I_C$ with
\begin{equation} \label{tr_CG_DG}
	I_C = \operatorname{tr}(\mathbf{C}).
\end{equation}
This quantity describes the length change of the object after a deformation.

\subsection{Deformation Energy}
The deformation energy or strain energy $\Psi$ is a scalar that defines the stored energy of a object undergoing a deformation. The strain energy density is the strain energy per unit volume. It equal to the work that has to be done by the stresses in order to alter the strains (\cite{KORSUNSKY20175}, p.10). This means that the energy is an indicator of how much force must be applied in order for an object to be deformed in a certain way. Thus we can use the deformation energy to express the relationship between the stresses and strains. We can illustrate this with a stress-strain curve shown in Fig. \ref{fig:stress_strain}. The area under the stress-strain curve corresponds to the strain energy.

\begin{figure}[!htbp]
	\centering
	\includegraphics[width=0.5\textwidth]{resources/stress_strain_curve.png}
	\caption[Stress-strain curve]{Example for a stress-strain curve\footnotemark}
	\label{fig:stress_strain}
\end{figure}
\footnotetext{https://commons.wikimedia.org/wiki/File:Stress-strain\_curve.svg}

Unfortunately, this relationship is not always linear, which makes it harder to find a suitable model. For example if we reach a stress state that produces yielding, we cannot assume that the relationship is linear. If the material reaches the yield point, it can no longer recover into its rest shape due to permanent fractures during the deformation. This point is illustrated in Fig. \ref{fig:stress_strain} with \textit{Yield stress}. \textit{UTS} in Fig. \ref{fig:stress_strain} is abbreviated for Ultimate tensile strength and defines the maximal stress an object can bear before breaking.

The behaviour of the object depends heavily on the material it consists of. This will be further explained in the next section concerning \textit{Material Constants}. In order to get a convincing simulation for a specific material, we have to choose an appropriate energy function. The behaviour of human flesh can be put into the category of hyperelastic materials. One key property of elasticity is, that if all the forces that are applied over an object add up to zero, the object must come back into its rest shape.

@@

In the context of flesh simulations, I am using the poisson's ratio as a characterization for the resistance to volume change of flesh. Biological tissues such as flesh, fat and muscles have one important property: Volume preservation. As a result the poisson's ratio takes on higher values in the range of 0.45 and 0.5 \cite{Smith:2018:SNF:3191713.3180491}.

The calculation of the poisson's ratio defined in Eq. \ref{eq:poisson} is a challenge. Fortunately, we can make use of the \textit{Lamé Parameters}, the two material specific constants $\mu$ and $\lambda$. With the help of these two constants we can transform Eq. \ref{eq:poisson} into the following form:

@@

This equation allows to calculate the poisson's ratio much easier.

>>>>

\chapter{Background} \label{c:Background}
My goal is the same as in the paper \textit{\acrshort{snh}}, namely to animate human-like characters. More precisely, the focus is on the behaviour of the character's flesh. This chapter serves as an introduction into the mathematical as well as physical background, needed to understand the upcoming calculations and conclusions. At the beginning of this chapter, I will define the notation and convention used throughout this whole thesis. Next, I will give insights in the mathematics used and present some of the concepts used in continuum mechanics. However, I will not include each proof explicitly, as there are already many good resources for an interested reader.

@@

The \acrlong{svd} (\acrshort{svd}) will play an important role in the formulation of the deformation gradient. It represents the best possible approximation of a given matrix by a matrix of low rank. This approximation can be looked at as a compression of the data given (\cite{LiesenMehrmann2015}, p. 295). Firstly we need to define what singular values are.

@@

This definition and theorem were taken from \textit{Numerical linear algebra with applications: Using MATLAB} (\cite{ford2014numerical}, p.113, p.300).

\subsection{Polar Decomposition}
Another theorem we will use in the next sections is the the Polar Decomposition Theorem.

@@

This theorem was taken from \textit{Continuum Mechanics} in which they include the proof for $3 \times 3$ matrices (\cite{Spencer1980}, p.12).

@@

We can imagine that we map each particle of a chosen object from its rest state to a deformed one. Each particle $X$ of a body can be characterized by a vector $\mathbf{x}$ containing its positional coordinates at time $t_0=0$. This is the reference configuration. If the particles are displaced after a time $t$, we can describe their new coordinate vector with 
\[
	\check{\mathbf{x}} = \phi (\mathbf{x}, t).
\]
If we consider, for example a \textit{Translation}, every particle is displaced by the same distance and direction. Hence, we can calculate $\check{\mathbf{x}}$ with
\[
	\check{\mathbf{x}} = \mathbf{x} + \mathbf{c}(t)
\]
where $\mathbf{c}$ is a vector that only depends on $t$ (\cite{Spencer1980}, p.63).

@@

\subsection{Deformation Gradient}
The next quantity I will use in the upcoming chapters is the deformation gradient \textbf{F}. With its help, we can calculate the change in volume and length of an object during a deformation. We can obtain \textbf{F} by taking the gradient of the function $\phi$ discussed in the previous section.

\begin{wrapfigure}[11]{r}{5.5cm}
\centering
	\includegraphics[width=5.5cm]{resources/stretch_plot.png}
	\caption[Stretching of a rectangle]{Stretching of a rectangle}
	\label{stretch:1}
\end{wrapfigure}
As an example, I am taking a simple deformation in 2D. The coloured area in Fig. \ref{stretch:1} represents an object in its rest state. We can stretch this area in the following way:
\begin{align*}
	\check{x} &= 1.5x + 0y		\\
	\check{y} &= 0x + 2y	
\end{align*}

The resulting deformation gradient can then be obtained by
\[
	\mathbf{F} = \begin{bmatrix} \frac{\partial [1.5x + 0y]}{\partial x} & \frac{\partial [1.5x + 0y]}{\partial y} \\ \frac{\partial [0x + 2y]}{\partial x} & \frac{\partial [0x + 2y]}{\partial y} \end{bmatrix} = \begin{bmatrix} 1.5 & 0.0 \\ 0.0 & 2.0 \end{bmatrix}.
\]
If \textbf{F} is equal to the identity matrix \textbf{I}, there is no deformation present. This would be the case for rigid body displacements. Since we will be handling deformation in the 3D space, the deformation gradient will have the form of a $3 \times 3$ matrix 
\begin{equation}\label{eq:deformation_gradient}
\textbf{F} = \left[ \,f_0\, \bigg| \,f_1\, \bigg| \,f_2\, \right] = \begin{bmatrix} f_0 & f_3 & f_6 \\ f_1 & f_4 & f_7 \\ f_2 & f_5 & f_8 \end{bmatrix}.
\end{equation}
\textbf{F} can be factorized and used to calculate other quantities. The next few sections will explain this.

@@

\subsubsection{Cauchy-Green}
The right Cauchy-Green tensor \textbf{C} can be calculated by 
\begin{equation}\label{CG_DG}
	\mathbf{C} = \mathbf{F^\intercal F}
\end{equation}
and is a $3 \times 3$ matrix for deformations in the 3D domain. Using \textbf{C}, we can calculate the first right Cauchy-Green invariant $I_C$ with
\begin{equation} \label{tr_CG_DG}
	I_C = \operatorname{tr}(\mathbf{C}).
\end{equation}
This quantity describes the length change of the object after a deformation.

\subsection{Deformation Energy}
The deformation energy or strain energy $\Psi$ is a scalar that defines the stored energy of a object undergoing a deformation. The strain energy density is the strain energy per unit volume. It equal to the work that has to be done by the stresses in order to alter the strains (\cite{KORSUNSKY20175}, p.10). This means that the energy is an indicator of how much force must be applied in order for an object to be deformed in a certain way. Thus we can use the deformation energy to express the relationship between the stresses and strains. We can illustrate this with a stress-strain curve shown in Fig. \ref{fig:stress_strain}. The area under the stress-strain curve corresponds to the strain energy. 
\begin{figure}[!htbp]
	\centering
	\includegraphics[width=0.5\textwidth]{resources/stress_strain_curve.png}
	\caption[Stress-strain curve]{Example for a stress-strain curve\footnotemark}
	\label{fig:stress_strain}
\end{figure}
\footnotetext{https://commons.wikimedia.org/wiki/File:Stress-strain\_curve.svg}

This relationship is not always linear. For example if we reach a stress state that produces yielding, we cannot assume that the relationship is linear anymore. If the material reaches the yield point, it can no longer recover into its rest shape due to permanent fractures during the deformation. This point is illustrated in Fig. \ref{fig:stress_strain} with \textit{Yield stress}. \textit{UTS} in Fig. \ref{fig:stress_strain} is abbreviated for Ultimate tensile strength and defines the maximal stress an object can bear before breaking.

The behaviour of the object depends heavily on the material it consists of. This will be further explained in the next section concerning \textit{Material Constants}. In order to get a convincing simulation for a specific material, we have to choose an appropriate energy function. The behaviour of human flesh can be put into the category of hyperelastic materials. The key property of elasticity is, that if all the forces that are applied over an object are removed, the object recovers into its original shape and volume (\cite{KORSUNSKY20175}, p. 5).

@@

In the context of flesh simulations, I am using the poisson's ratio as a characterization for the resistance to volume change of flesh. Biological tissues such as flesh, fat and muscles have one important property: Volume preservation. As a result the poisson's ratio takes on higher values in the range of 0.45 and 0.5 (\cite{Smith:2018:SNF:3191713.3180491}, 12:1).

The calculation of the poisson's ratio defined in Eq. \ref{eq:poisson} is a challenge. Fortunately, we can make use of the \textit{Lamé Parameters}, the two material specific constants $\mu$ and $\lambda$. With the help of these two constants we can transform Eq. \ref{eq:poisson} into the form

@@

This equation allows to calculate the poisson's ratio much easier (\cite{BERGSTROM2015209}, p. 231).

### 2020-04-04 LaTeX-Master-Vorlage/chapter/Introduction.tex
background
working on paper

\section{Motivation}
With steadily increasing computational power, the demand of better results is constantly growing. Especially in the field of animation and simulation we are no longer happy with mediocre results. 
In the entertainment sector and gaming industry animation studios like Pixar\textsuperscript{\textcopyright} or Disney\textsuperscript{\textcopyright} brought us games and movies of highest quality. Both of them have made groundbreaking progress over the years. This is easily observed when we compare today's work with that from ten years ago.

@@

In the paper \textit{\acrlong{snh}} \cite{Smith:2018:SNF:3191713.3180491}, the authors addressed exactly this problem of making an animated movement of a human-like character look as natural as possible. Consequently this thesis is based heavily on this paper. In the following, I will abbreviate the name of the paper with \textit{\acrshort{snh}}. \\
But before diving further into the content of the paper, a fundamental background is needed. In order to animate a physical movement, we first need to understand the physics behind it which requires some knowledge in the field of continuum mechanics. Unfortunately, for most of us it has yet to be learned. The goal of this thesis is to give the necessary physical and mathematical background for a regular computer science student to understand the field of animation. In addition I will go deeper into the thematics of the paper \textit{\acrshort {snh}}. I aim to get an understanding of their contribution in the field and implement their proposed energy myself.

\section{Structure}
In the following, I will give a brief overview of the necessary mathematical background and deliver an introduction into the field of continuum mechanics. Next, I will go through the ideas mentioned in the paper and include some calculations and visualisations that serve for a better understanding. Lastly I will give an insight into the process of implementing the energy. \\

\todoredefined[inline]{
TODO: Adjust according to additions in text. Maybe add other applications as well (physical simulation) and how to go over to it?
}

>>>>

\section{Motivation}
With steadily increasing computational power, the demand of better looking results is constantly growing. Especially in the field of animation and simulation we are no longer happy with mediocre results. 
In the entertainment sector animation studios like Pixar\textsuperscript{\textcopyright} brought us movies of highest quality. They have made groundbreaking progress over the years. This can easily be observed when we compare today's work with that from ten years ago.

@@

In the paper \textit{\acrlong{snh}} \cite{Smith:2018:SNF:3191713.3180491}, the authors addressed exactly this problem of making an animated movement of a human-like character look as natural as possible. Consequently, this thesis is based heavily on this paper. In the following, I will abbreviate the name of the paper with \textit{\acrshort{snh}}. \\
But before diving further into the content of the paper, a fundamental background is needed. In order to animate a physical movement, we first need to understand the physics behind it. This requires some knowledge in the field of continuum mechanics. Unfortunately, for most of us it has yet to be learned. The goal of this thesis is to give the necessary physical and mathematical background for a regular computer science student to understand the field of animation. In addition I will go deeper into the thematics of the paper \textit{\acrshort {snh}}. I aim to get an understanding of their contribution in the field and implement their proposed energy myself.

\section{Structure}
In the following, I will give a brief overview of the necessary mathematical background and deliver an introduction into the field of continuum mechanics. Next, I will go through the ideas mentioned in the paper and include some calculations and visualisations that serve for a better understanding. Lastly I will give an insight into the process of implementing the energy.

### 2020-04-04 LaTeX-Master-Vorlage/chapter/paper.tex
working on paper

\section{Deformation Gradient}
In the following calculations some properties of the deformation gradient $\mathbf{F}$ covered in chapter 2 will be used. These properties are summarized in table \ref{table:gradient_quantities} for a better overview:

@@

\section{Energy Formulation}
\subsection{Stability}
The core goal of the paper was to model deformations for virtual characters that have human-like features. The focus of the authors was the deformation energy. In order to achieve a convincing result the first step is to specify some requirements. Firstly the energy has to be stable. More precisely, we need a hyperelastic energy that is stable in the following four important ways:

>>>>

\section{Deformation Gradient}
In the following calculations the properties of the deformation gradient $\mathbf{F}$ covered in chapter 2 will be used. These properties are summarized in table \ref{table:gradient_quantities} for a better overview.

@@

\section{Energy Formulation}
\subsection{Stability}
The core goal of the paper was to model deformations for virtual characters that have human-like features. In order to achieve better results than what has been done in current research, they formulated a new deformation energy. In chapter 2 I concluded that the appropriate energy for animating soft tissues as flesh has to be hyperelastic. An important property is the stability of the energy. We need an energy that is stable in the following four ways:
